Evaluation of Pre-Trained Models for Translation in Different Languages

### 2024-08-03 paper/paper.tex
nllb done

Pre-trained models are often not evaluated under the same datasets or benchmarks \cite{liu-2020-mbart,nllb200-2020,fan-2020-m2m100,wei-2023-polylm}. Therefore, the performance of these models across different languages seem to be yet unclear.

@@

@@

@@

\begin{comment}
PolyLM \cite{wei-2023-polylm} (32 languages), and
mT5 \cite{xue-2021-mt5} (101 languages).

\subsection{Sequence-to-Sequence Models}

GPT \cite{openai-2024-gpt4}

mBERT \cite{devlin-2019-bert} is the multilingual version of BERT, introduced in the same paper, trained on 104 different languages. However, the model been shown to suffer on low-resource languages \cite{wu-2020-mbert-are-all}.

T5 \cite{raffel-2023-t5} is a Text-to-Text Transfer Transformer model trained on English.

MistralAI \cite{jiang-2023-mistral-7b}

\end{comment}

>>>>

many-to-one? ... text from multiple source languages into a single target language (so this is many-to-one)

(Using many-to-many models for many-to-one translation is a practical approach, as these models are designed to handle multiple languages and can efficiently translate between any pair of supported languages. Many-to-many models leverage shared multilingual representations, making them well-suited for translating text from various source languages into a single target language.)

Despite these advancements, pre-trained models are often evaluated using different datasets and benchmarks \cite{liu-2020-mbart,nllb200-2020,fan-2020-m2m100,wei-2023-polylm}, making it difficult to gauge their relative effectiveness across various languages.

Only works by META AI are benchmarked thoroughly with popular datasets across various sources.

There does not seem to be any works done on comparing or benchmarking pre-trained models in machine translation.

@@

(Multilinguality has been explored in the supervised NMT literature, where it has been shown to enable information sharing among related languages.) \cite{garcia-2020-multilingual}

(NMT systems have lower translation quality on very long sentences) \cite{koehn-2017-challenges}
(they often require large amounts of parallel (source, target) sentence pairs to train) \cite{koehn-2017-challenges}

Despite fine-tuning pre-trained multilingual models shows improved performance... \cite{cooper-stickland-2021-recipes}

% Relevance and Impact of Machine Translation:
% Discuss the broader implications of accurate machine translation beyond just convenience, such as its impact on global communication, business, and cross-cultural exchange.

% Challenges and Limitations:
% Briefly mention some common challenges in machine translation, such as handling idiomatic expressions, preserving context, and managing low-resource languages.

% Novelty and Contributions:
% Emphasize what your paper contributes to the existing body of research. For instance, is your evaluation filling a gap in the current literature? Are there specific insights or implications for future research or practical applications?

% Paper Organization:
% Briefly outline the structure of the paper to guide the reader on what to expect in subsequent sections.

@@

\subsection{Evaluation Dataset}

IWSLT \cite{agarwal-iwstl-2023}

TED \cite{ye-2018-word-embeddings-ted}

(Flores-101 evaluation benchmark, consisting of 3001 sentences extracted from English Wikipedia and covering a variety of different topics and domains. These sentences have been translated in 101 languages by professional translators through a carefully controlled process. The resulting dataset enables better assessment of model quality on the long tail of low-resource languages, including the evaluation of many-to-many multilingual translation systems, as all translations are fully aligned) \cite{goyal-2022-flores}

@@

mRASP2 \cite{pan-2021-mrasp2}

\subsection{Evaluation Metrics}

% Introduce the concept of evaluation metrics for MT, explaining why BLEU is chosen for this study and if there are other metrics that could complement it (like METEOR or ROUGE).

% \begin{comment}
% PolyLM \cite{wei-2023-polylm} (32 languages), and
% mT5 \cite{xue-2021-mt5} (101 languages).

% \subsection{Sequence-to-Sequence Models}

% GPT \cite{openai-2024-gpt4}

% mBERT \cite{devlin-2019-bert} is the multilingual version of BERT, introduced in the same paper, trained on 104 different languages. However, the model been shown to suffer on low-resource languages \cite{wu-2020-mbert-are-all}.

% T5 \cite{raffel-2023-t5} is a Text-to-Text Transfer Transformer model trained on English.

% MistralAI \cite{jiang-2023-mistral-7b}

% \end{comment}
